Exception Handling Bug Hazards in Android

### 2015-10-06 SPRINGER_JOURNAL/msr_journal.tex
survey section

\subsection{Replication Package}
All the data used in this study is publicly available at tool website hosted in GitHub: https://github.com/souzacoelho/exceptionminer.
Specifically we provide: (i) all issues related to Android projects found
in GiTHub and Google Code used in this study; (ii) all stack traces extracted
from issues; (iii) the results of manual inspection steps; (iv) the
ExceptionMiner tool we developed to support stack trace extraction and distilling.

@@

When developing Java-based applications it is inevitable to deal with exceptions. Consequently, our first question explores how often developers deal with the exception handling code in Android development:

\emph{\textbf{RQ 1:} How often developers deal with the exception handling in Android development?} To make the analysis easier, we further refine RQ1 in the following subquestions:
\begin{itemize}
\item RQ1.1:\emph{How often developers handle exceptions?}
\item RQ1.2:\emph{How often developers throw exceptions?}
\end{itemize}

Another issue that show us how developers may deal with such bug hazards is
 their knowledge about exception handling best practices presented in Section~\ref{sec:best}.
 This leads us to our second research question:

\emph{\textbf{RQ 2:} Do Android developers adopt exception handling best practices?}

\emph{\textbf{RQ 4:}  Do Android developers consider implementing the exception handling code of Android apps more difficult than general Java applications?}

Then the third research question focus on how developers deal with the main bug hazards found during the mining study.

\emph{\textbf{RQ 3:}What are the developers perspectives about the exception handling bug hazards?} To answer such question we further refine RQ3 in the following subquestions:
\begin{itemize}
\item\emph{RQ3.1:Have developers ever faced a crash caused by NullPointerException?}
\item\emph{RQ3.2:Are developers aware of the robustness threats caused by cross-type wrappings?}
\item\emph{RQ3.3:Are developers aware of the robustness threats caused by JNL non-documented checked exceptions?}
\end{itemize}

@@

@@

\noindent\emph{\textbf{Q2: How the exception handling code affect the development of robust apps and how developers prevent apps from crashing?}}

@@

Developers were also asked whether the exception handling code helped with the development of robust Android applications. Although most of the developers answered positively (78.9\%) when asked to provide one or more reasons for their answers, most of the respondents also provide reasons why the exception handling code can sometimes impairs the robustness.  Since positive and negative reasons were provided intermixed we analyzed both set of answers and ranked the top reasons why exception handling that can negatively or positively affect robustness according to developers perceptions.

 Table ~\ref{tab:topreasons} presents the top reasons why the exception handling code may or may not help de development of robust apps. Although most of the respondents answered that exception handling code helped with the development of robust Android applications, almost every respondent pointed a drawback associated to he exception handling code, saying that if not used with care the exception handling code may lead to app crashes.

//// research

\noindent\emph{\textbf{Q3.1: How NullPointerException impact the development of robust Android apps?}}

//////NULLPONTER WORSE IN ANDROID /////
Developers were also asked whether or not craches caused by NullPointerException were most common in Android development than in stantard Java development.  Almost half of them (XXX\%) mentioned that Android apps were more suceptible to NullPointerException crashes. Some of the reasons mentioned by developers were: (i) the activity and fragment lifecycle - since both are constantly distroyed its references can frequently become null leading to NullPointerExceptions; (ii) the framework complexity - several methods od Android framework return null, and there is not enought documentation to inform the developer when using such methods. Moreover, few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table X. The other half mentoned that NullPointerExceptions are  common in Android development as it is common in Java standard development.		//////

@@

\greybox{This is nothing specific to Android, it's just how Java works.}

@@

////// PREVENT NULL POINTER /////
Developers were also asked about what they usually do to prevent NullPointerExceptions to happen when their app is in production. Most of them sais that include null-checks on the application code (59\%) mostly when checking the return of a method and as method guard conditions. Many respondents (39\%) also answered that instead of just sprinkling the code with null-checks when a NullPointerException happens they investigate the cause of the null reference and work to fix it, instead of just adding a null check. Some respondents also mention that use annotations such as @Nullable @NonNull, provided in Android since version XXX; such annotations guide de developer about what can and cannot be null during the application lifecycle. Altough catching the NullPointerExceptions is considered a bad practice some respondents said that they do it to prevent crashes caused by NullPointerExceptions. Such behavior was mentioned by one of the responded as the behavior adopted as as way non-expericente developers take to prevent such crashes - which is undoubtly a mistake. Some respondents also mentioned to replace the NullPointerException by IllegalArgumentException when a method should not receive a null as parameter. This is a good exception handling practice mentioned by Bloch in [ref]. Only few developers mentioned about coding standards and static code analyzers to prevent it (2\%).////////

@@

\noindent\emph{\textbf{Q3.2: How Cross-Type Wrappings impact the development of robust Android apps?}}

@@

Some of the best practices are related to the different ways of handling checked and unchecked exceptions. When the designer of an API specifies that a method throws a checked exception, it is telling
the caller of the API that such exception should be handled.

///// HANDLING RUNTIME AND CHECKED /////
Developers were presented to two pieces of code: one on which a RuntimeException was caught and another one where an IOException has caught; both in the contexts of Activity life-cycle methods. Then de developers were asked whether or not the way to handle a runtime exception should be different from the way to handle a checked exception.

This question was motivated by the fact that during the mining study we could identify that many checked exception were just wrapped in a runtime exception and rethrown, as a way to by pass any kind of handling. When asked about of how to deal with a checked exception signaled by a method (invoked in the context of onPause() activity method), most respondents (62.50\%) said that they should add a try-catch block surrounding the method invocation. Also most of the respondents said that to deal with a method that is signaling a runtime exception we should surround such method by a try-catch block (xxxxx\%). However, what developers mentioned as handling actions (what is performed inside the catch clause to deal with the exception) differes in both cases. One of them mentioned that when the designer of an API decides to signal a checked exception s/he is telling the developer that the exception should be handled. Hence, some handling actions mentioned for checked exceptions were: (i) retrying the same operation; (ii) involving the user in finding a solution to the exception condition such as opening a pop-up and asking the user to define a new place to save the file (as in the example presented to them involved a IOException been signaled); and also (iii) presenting an error message to the user. On the other hand, the most common way of dealing with a runtime exception would be to catch it and present an error message to the user, or swallow the exception. Some developers also mentioned to use a Toast to support this task - a toast is an Android component that provides simple feedback about an operation in a small popup. (http://developer.android.com/guide/topics/ui/notifiers/toasts.html). Few developers mentioned that the runtime exception should flow upstream so the framework so it could gracefully crash the app. Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, most Android developers condireded checked exceptions a way od using exceptions that can prevent uncaught exception crashes - since in order to a checked exception to flow upstream and crash the app the developers needs to explicitly do so. //////

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
\hline
add a try-catch block &	37 &	59\% \\
present error message	&  14 &	22\% \\
log the exception	& 12 &	19\% \\
throw a checked exception	& 12 & 19\% \\
let it crash / crash fast 	& 6 &	10\% \\
swallow the exception	& 5 &	8\% \\
add a throws declaration	& 5 &	8\% \\
retry &	5 &	8\% \\
let it crash &	4 &	6\% \\
report crash &	3 &	5\% \\
Use Toast & 	3 &	5\% \\
\hline
\end{tabular}
\caption{Top ways of handling runtime exceptions -  63 non-empty responses. }
\label{tab:handlingruntime}
\end{table}

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
\hline
add a try-catch block	& 25 &	63\% \\
present error message	& 7 &	18\% \\
involve the user in solving	& 5 &	13\% \\
retry  &	5 &	13\% \\
investigate the cause &	3 &	8\% \\
same as runtime	& 3 &	8\% \\
add a throws declaration / pass upstream &	2 &	5\% \\
swallow	& 1 &	3\% \\
toast	 & 1 & 3\% \\
\hline
\end{tabular}
\caption{Top ways of dealing with a checked exception signaled in the context of an Activity method -  40 non-empty responses. }
\label{tab:handlingruntime}
\end{table}

/////// CROSS TYPE WRAPPING //////
Developers were presented to a code snipet on which as cross-type wrapping was performed - as illustrated in Figure X. Then developers were asked whether such cross-type wrapping could affect the application robustness in some way. 18\% of the respondents mentioned that the app would crash anyway, regardless the exception was wrapped or not. Most of the respondents, however, mentioned disadvantages of such wrapping such as: (i) it impairs proper handling since when wrapping in a general exception we loose information about the exception situation to he handled (24\%); such wrapping treat all exceptions as critical in other words, every exception will remain uncaught and will lead to an app crash; it does not allow the developer to perform a proper handling (e.g. retry) for exception types that are not critical. 		//////

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
\hline

impairs proper handling (loses exception information)	& 12 &	24\% \\
uncaught will crash the app	& 12 &	24\% \\
app will crash anyway	& 9	& 18\% \\
should catch / handle properly (do local recovery)	 & 5 & 	10\% \\
treat all exceptions as critical	& 5 &	10\% \\
useless rethrow	& 2 &	4\% \\
activity mehtods cannot throw exceptions &	1 &	2\% \\

\hline
\end{tabular}
\caption{Top reasons why cross-type wrapping may impact app robustness - 49 non-empty responses. }
\label{tab:handlingruntime}
\end{table}

\bigskip
\noindent\emph{\textbf{RQ 3.3 Are developers aware the hidden dangers of non-declared checked exceptions?}}
\bigskip

//////// JNI /////
Only 3 respondents out of 71 (4\% of respondents) were aware of the fact that there are ways a method can throw a non-declared checked exception such as JNI/native code and relfection or directly changing the bytecode/dalvik code bypassing the compile-time checking.  Most of the repondents were not aware that a method could throw a checked exception withou declaring it on its signature, regardless the circunstance. Most respondents mentioned that if a method throws a non-declared exception it must be a RuntimeException. We could observe that this exception handling bug hazard, although represent a threat to app robustness, is hardly known by the Android developers involved in this survey, and can represent.	////

@@

\noindent\emph{\textbf{Q2: Do developers know about exception handling best practices?}}

We can observe that 62\% of surveyed Android developers follow Java exception handling best practices. Among the best practices cited by developers the most popular ones were: avoid generic catches and avoid empty catch blocks. They did not mention a any guide or reference of best practices. When asked about best practices specific to Android developers, most of the developers mentioned that they did no about them. The ones who answered positively said that they adopt the same best practices as in any Java program.

@@

\noindent\emph{\textbf{Q3:What are the developers perspectives about the exception handling bug hazards?}}

\greybox{Sometimes when you develop multi-threaded applications, when the thread finishes you don't have any GUI (or Context) to show information, so you have to be aware of your application's state (foreground, background)}

Handle life cycle of activities and fragments.

Q10: What makes exception handling in Android different from exception handling in Java?

The Android framework is what adds the complexity in figuring out what caused an exception because more often than not, the error is triggered from the framework as a result of something else you did

Sometimes when you develop multi-threaded applications, when the thread finishes you don't have any GUI (or Context) to show information, so you have to be aware of your application's state (foreground, background).

Application-Activity-Fragment lifecycle.

You can wrap the main function in java in try/catch in android you can\'t but there are equivalents

Though not particular to Android, most Android apps have several different entry points: Activities, Services, etc, can be started by different services; same thing with events. It may be hard to ensure that all this cases are properly handled. The logcat is a helpful way of tracking exceptions when developing and testing on real phones.

Any uncatched exception will crash the app. There are many unknown exception thrown by the system, that are only happen once in a lifetime like IllegalStateException: eglMakeCurrent failed EGL\_BAD\_CONTEXT

The lifecycle of activities\/fragements etc make it harder to work out what order things are called in and so exceptions can occur because you didn't realise that another method isn't called

@@

\begin{quotation}
\noindent
 \emph{``Everybody hates thinking about exceptions, because they are not supposed to happen''\\
  (Brian Foote)}\footnote{Brian Foote shared his opinion in a conversation with James Noble - quoted on the paper: hillside.net/plop/2008/papers/ACMVersions/coelho.pdf}
\end{quotation}

\emph{\textbf{The exception handling confusion problem.}}
When (mis)applied, exception wrapping can make the exception-related code
 more complex and lead to what we call the \emph{exception handling confusion problem}.
This problem can lead the program to an unpredictable state in the presence of exceptions,
as illustrated by the scenario in which a checked exception wraps an OutOfMemoryError. 
Currently there is no way of enforcing Java exception type conventions during program development.
Hence, further investigation is needed on finding ways to help developers in dealing with this
 problem, either preventing odd wrappings or enabling the developer to
 better deal with them.
Furthermore, this calls for empirical studies on the actual usefulness of Java's hybrid exception model.

@@

\emph{\textbf{On the null pointer problem.}}
The null reference was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake''~\cite{hoare2}.
In this sudy, the null references were, infact, responsible for several reported issues - providing further evidence to Hoare's statement.
This observation emphasizes the need for solutions to avoid NullPointerExceptions, such as:
(i) lightweight intra-method null pointer analysis as supported by
Java 8 @Nullable annotations\footnote{Already supported by tools such as Eclipse, IntelliJ, Android Studio 0.5.5 (release Apr. 2014) to detect potential 
null pointer dereferences at compile time.};
(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha~\cite{nanda2009accurate};
or (iii) language designs which avoid null pointers, such 
as Monads~\cite{Walde95} (as used in functional languages for values that may not be available 
or computations that may fail) could improve the robustness of Java programs.

\emph{\textbf{Preventing uncaught exceptions}}
In this study we could observe undocumented runtime exceptions thrown by third party code,
and even undocumented checked exception thrown by a JNI interface.
Such undocumented exceptions make it difficult, and most of the times infeasible
for the client code to protect against `unforeseen'' situations that may happen 
while calling a library code.

One may think that the solution for the uncaught exceptions may be to define a general handler, 
which is responsible for handling any exception that is not
adequately handled inside the applications. Although this 
solution may prevent  the system from abruptly crashing,
 such a general handler will not have enough
contextual information to adequately handle the exception, 
beyond storing a message in a log file and restarting the application.
However, such a handler cannot replace a carefully designed exception 
handling policy~\cite{Robil00}, which requires
third-party documentation on the exceptions that
may the thrown by APIs used. 
Since  documenting runtime exceptions is a tedious and error prone
task, 
this calls for tool support to automate the extraction of runtime exceptions
from library code. Initial steps in this direction have been proposed
by van Doorn and Steegmans \cite{van2005combining}.

\section{Threats to Validity}
\label{sec:threats}

%\noindent
\emph{Internal Validity.} We used a heuristics-based parser to mine
exceptions from issues.  Our parsing strategy was conservative by default; for
example, we only considered exception names using a fully qualified class name
as valid exception identifiers, while, in many cases, developers use the
exception name in issue description. Conservative parsing may minimize false
positives, which was our initial target, but also tends to increase false
negatives, which means that some cases may have not been identified as
exceptions or stack traces. Our limited manual inspection did not reveal such
cases. Moreover, in this study we manually mapped the concerns related
to exceptions. To ensure the quality of the analysis, we calculated the
 interrater agreement after three independent 
developers classified a randomly selected sample (of 25 exception 
types from the total of 100); the interater agreement was high (96\%).

@@

Another threat relates to the fact that parts of our analysis 
are based on the availability of stack traces on issues reported on Github and Googlecode projects. 
In using these datasets, we make an underlying assumption: the stack traces reported on issues are 
representative of valid crash information of the applications. 
One way to mitigate this threat would be to access to the full 
set of crash data per application. Although some services exist 
to collect crash data from mobile applications~\cite{BugSe14,BugSn14,Googl14,Acra14},
they do not provide open access to the crash reports of their client applications.
In our study, we mitigated this threat by manually inspecting
the source code associated to a subset of the reported exception stack traces.
This subset comprises the stack traces related to the main findings 
of the study (e.g., ``undocumented runtime and checked exceptions'',
and ``cross-type wrappings'').
%\noindent

>>>>

@@

When developing Java-based applications it is inevitable to deal with exceptions. Consequently, our first question explores how developers deal with the exception handling code in Android development:

\emph{\textbf{RQ 1:} How often developers with the exception handling code while developing Android apps?}

To make the analysis easier, we further refine RQ1 in subquestions, such as: \emph{How often developers handle exceptions?}; \emph{How often developers throw exceptions?}. Moreover, another question that show us how developers may deal with the EH code is: \emph{Do they knew about Java EH best practices and/or Android specific EH best practices?}

Then following research question focus on the developers perspectives about the main bug hazards found during the mining study, more specifically:

\emph{\textbf{RQ 2:} How the NullPointerExceptions impact the development of robust Android apps?}

\emph{\textbf{RQ 3:} How cross-type wrappings impact the development of robust Android apps?}

\emph{\textbf{RQ 4:} Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}

%\emph{\textbf{RQ 4:}  Are developers aware of the non-declared checked exceptions?}

%\emph{RQ 2:Have developers ever faced a crash caused by NullPointerException?}
%\emph{RQ 3:Are developers aware of the robustness threats caused by cross-type wrappings?}; %\emph{RQ 4: Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}

@@

Finally developers were asked whether the exception handling code helped with the development of robust Android applications, and what they usually do to prevent apps from crashing. Hence, the last research question that guided this survey was the following:

\emph{\textbf{RQ 5:} How the exception handling code affect the development of robust apps and how developers prevent apps from crashing?}

@@

\noindent\emph{\textbf{Q 2: How the NullPointerExceptions impact the development of robust Android apps?}}

@@

%//////NULLPONTER WORSE IN ANDROID /////
Developers were also asked whether or not craches caused by NullPointerException were most common in Android development than in stantard Java development.  Almost half of them (XXX\%) mentioned that Android apps were more suceptible to NullPointerException crashes. Some of the reasons mentioned by developers were: (i) the activity and fragment lifecycle - since both are constantly distroyed its references can frequently become null leading to NullPointerExceptions; (ii) the framework complexity - several methods od Android framework return null, and there is not enought documentation to inform the developer when using such methods.

\greybox{The Android framework is what adds the complexity in figuring out what caused an exception because more often than not, the error is triggered from the framework as a result of something else you did.}

Moreover, few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table X. The other half mentoned that NullPointerExceptions are  common in Android development as it is common in Java standard development.

@@

%\greybox{This is nothing specific to Android, it's just how Java works.}

@@

%////// PREVENT NULL POINTER /////
Developers were also asked about what they usually do to prevent NullPointerExceptions to happen when their app is in production. Most of them sais that include null-checks on the application code (59\%) mostly when checking the return of a method and as method guard conditions. Many respondents (39\%) also answered that instead of just sprinkling the code with null-checks when a NullPointerException happens they investigate the cause of the null reference and work to fix it, instead of just adding a null check. Some respondents also mention that use annotations such as @Nullable @NonNull, provided in Android since version XXX; such annotations guide de developer about what can and cannot be null during the application lifecycle. Altough catching the NullPointerExceptions is considered a bad practice some respondents said that they do it to prevent crashes caused by NullPointerExceptions. Such behavior was mentioned by one of the responded as the behavior adopted as as way non-expericente developers take to prevent such crashes - which is undoubtly a mistake. Some respondents also mentioned to replace the NullPointerException by IllegalArgumentException when a method should not receive a null as parameter. This is a good exception handling practice mentioned by Bloch in [ref]. Only few developers mentioned about coding standards and static code analyzers to prevent it (2\%).

@@

\noindent\emph{\textbf{RQ 3: How Cross-Type Wrappings impact the development of robust Android apps?}}

@@

Some of the best practices are related to the different ways exceptions should be handled according to whether it is a checked and an unchecked exception. When the designer of an API specifies that a method throws a checked exception, it is telling the caller of the API that such exception should be handled.

%///// HANDLING RUNTIME AND CHECKED /////		
To acess developers perspective on whether checked and unchecked exceptions should be handled different inside the app, the developers were presented to two pieces of code: one on which a RuntimeException was caught and another one where an IOException has caught; both in the contexts of Activity life-cycle methods. Then the developers were asked whether or not the way to handle a runtime exception should be different from the way to handle a checked exception.

This question was motivated by the fact that during the mining study we could identify that many checked exception were just wrapped in a runtime exception and rethrown, as a way to by pass any kind of handling. When asked about of how to deal with a checked exception signaled by a method (invoked in the context of onPause() activity method), most respondents (62.50\%) said that they should add a try-catch block surrounding the method invocation. Also most of the respondents said that to deal with a method that is signaling a runtime exception we should surround such method by a try-catch block (xxxxx\%). However, what developers mentioned as handling actions (what is performed inside the catch clause to deal with the exception) differes in both cases.

One of the respondents mentioned that when the designer of an API decides to signal a checked exception s/he is telling the developer that the exception should be handled. Hence, some handling actions mentioned for checked exceptions were: (i) retrying the same operation; (ii) involving the user in finding a solution to the exception condition such as opening a pop-up and asking the user to define a new place to save the file (as in the example presented to them involved a IOException been signaled); and also (iii) presenting an error message to the user.

On the other hand, the most common way of dealing with a runtime exception would be to catch it and present an error message to the user, or swallow the exception. Some developers also mentioned to use a Toast to support this task\footnote{A toast is an Android component that provides simple feedback about an operation in a small popup - http://developer.android.com/guide/topics/ui/notifiers/toasts.html}. Few developers mentioned that the runtime exception should flow upstream so the framework so it could gracefully crash the app. Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, most Android developers condireded checked exceptions a way od using exceptions that can prevent uncaught exception crashes - since in order to a checked exception to flow upstream and crash the app the developers needs to explicitly do so.

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
\hline
Add a try-catch block &	37 &	59\% \\
Present error message	&  14 &	22\% \\
Log the exception	& 12 &	19\% \\
Throw a checked exception	& 12 & 19\% \\
Let it crash / crash fast 	& 6 &	10\% \\
Swallow the exception	& 5 &	8\% \\
Add a throws declaration	& 5 &	8\% \\
Report crash &	3 &	5\% \\
Use Toast & 	3 &	5\% \\
\hline
\end{tabular}
\caption{Top ways of handling runtime exceptions -  63 non-empty responses. }
\label{tab:handlingruntime}
\end{table}

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
\hline
Add a try-catch block	& 25 &	63\% \\
Present error message	& 7 &	18\% \\
Involve the user in solving	& 5 &	13\% \\
Retry  &	5 &	13\% \\
Investigate the cause &	3 &	8\% \\
Same way as runtime	& 3 &	8\% \\
Add a throws declaration / pass upstream &	2 &	5\% \\
%Swallow	& 1 &	3\% \\
%Toast	 & 1 & 3\% \\
\hline
\end{tabular}
\caption{Top ways of dealing with a checked exception signaled in the context of an Activity method -  40 non-empty responses. }
\label{tab:handlingruntime}
\end{table}

%/////// CROSS TYPE WRAPPING //////
Hence, developers were presented to a code snipet on which as cross-type wrapping was performed\footnote{This cross-type wrapping was found in several applicarions during the mining study as well as in some classes of Android framework} - as illustrated in Figure X. Then developers were asked whether such cross-type wrapping could affect the application robustness in some way. Table XXXX illustrates the  most mentioned reasons why cross-type wrapping may affect app robustness; 18\% of the respondents mentioned that the app would crash anyway, regardless the exception was wrapped or not. Most of the respondents, however, mentioned disadvantages of such wrapping such as: (i) it impairs proper handling since when wrapping in a general exception we loose information about the exception situation to he handled (24\%); such wrapping treat all exceptions as critical in other words, every exception will remain uncaught and will lead to an app crash; it does not allow the developer to perform a proper handling (e.g. retry) for exception types that are not critical.

{\footnotesize
\begin{verbatim}
  @Override
  public void onPause() {
     try {
        ...
      } catch (Exception e) {
            throw new RuntimeException("...", e);
      }
 }
\end{verbatim}
}

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top Reasons Why Cross-Type Wrapping Affect Robustness} & \bfseries{\#} & \bfseries{\%} \\
\hline
impairs proper handling (loses exception information)	& 12 &	24\% \\
uncaught will crash the app	& 12 &	24\% \\
app will crash anyway	& 9	& 18\% \\
should catch / handle properly (do local recovery)	 & 5 & 	10\% \\
treat all exceptions as critical	& 5 &	10\% \\
useless rethrow	& 2 &	4\% \\
activity mehtods cannot throw exceptions &	1 &	2\% \\

\hline
\end{tabular}
\caption{Top reasons why cross-type wrapping may affect app robustness - 49 non-empty responses. }
\label{tab:handlingruntime}
\end{table}

\bigskip
\noindent\emph{\textbf{RQ 4 Are developers aware of the non-declared checked exceptions?}}
\bigskip

%//////// JNI /////
The respondents were presented to a piece of code on which a non-declared checked exception has scaped, and were asked if they new any reason for it to happen. Only 3 respondents out of 71 (4\% of respondents) were aware of the fact that there are ways a method can throw a non-declared checked exception, such as JNI/native code, relfection, and directly changing the bytecode/dalvik code which bypass the compile-time checking.  Most of the repondents were not aware that a method could throw a checked exception withou declaring it on the method's signature. Most respondents mentioned that if a method throws a non-declared exception it must be a RuntimeException. We could observe that this exception handling bug hazard detected during the mining study, although represent a threat to app robustness, is hardly known by the Android developers involved in this survey. Below we can find a quote from one of the 3 respondents aware of JNI non-delcared checked exceptions.

@@

%\noindent\emph{\textbf{Q2: Do developers know about exception handling best practices?}}
%We can observe that 62\% of surveyed Android developers follow Java exception handling best 
%practices. Among the best practices cited by developers the most popular ones were: avoid generic 
%catches and avoid empty catch blocks. They did not mention a any guide or reference of best practices. 
%When asked about best practices specific to Android developers, most of the developers mentioned that 
%they did no about them. The ones who answered positively said that they adopt the same best practices 
%as in any Java program.

\noindent\emph{\textbf{RQ5: How the exception handling code affect the development of robust apps and how developers prevent apps from crashing?}}

@@

Developers were also asked whether the exception handling code helped with the development of robust Android applications. Although most of the developers answered positively (78.9\%) when asked to provide one or more reasons for their answers, most of the respondents also provide reasons why the exception handling code can sometimes impairs the robustness.  Since positive and negative reasons were provided intermixed we analyzed both set of answers and ranked the top reasons why exception handling that can negatively or positively affect robustness according to developers perceptions. Table ~\ref{tab:topreasons} presents the top reasons why the exception handling code may or may not help de development of robust apps.

Although most of the respondents answered that exception handling code helped with the development of robust Android applications, almost every respondent pointed a drawback associated to he exception handling code, saying that if not used with care the exception handling code may lead to app crashes.

\greybox{Any uncatched exception will crash the app. There are many unknown exception thrown by the system, that are only happen once in a lifetime like IllegalStateException: eglMakeCurrent failed EGL\_BAD\_CONTEXT}

\greybox{The lifecycle of activities\/fragements etc make it harder to work out what order things are called in and so exceptions can occur because you didn't realise that another method isn't called.}

\greybox{Though not particular to Android, most Android apps have several different entry points: Activities, Services, etc, can be started by different services; same thing with events. It may be hard to ensure that all this cases are properly handled. }

\greybox{You can wrap the main function in java in try/catch in android you can\'t. }

\section{Replication Package}
All the data used in the mining study and in the survey-based study is publicly available at the ExceptionMiner tool website hosted in GitHub\footnote{ https://github.com/souzacoelho/exceptionminer}.
Specifically we provide: (i) all issues related to Android projects found
in GiTHub and Google Code used in this study; (ii) all stack traces extracted
from issues; (iii) the results of manual inspection steps; (iv) the
ExceptionMiner tool we developed to support stack trace extraction and distilling; (v) the survey questions; and (iv) the survey responses and summary.

@@

\begin{quotation}
\noindent
 \emph{``Everybody hates thinking about exceptions, because they are not supposed to happen'' 
  (Brian Foote)}\footnote{Brian Foote shared his opinion in a conversation with James Noble - quoted on the paper: hillside.net/plop/2008/papers/ACMVersions/coelho.pdf}
\end{quotation}

%\emph{\textbf{The exception handling confusion problem.}}

\subsection{The exception handling confusion problem}
When (mis)applied, exception wrapping can make the exception-related code
 more complex and lead to what we call the \emph{exception handling confusion problem}.
This problem can lead the program to an unpredictable state in the presence of exceptions,
as illustrated by the scenario in which a checked exception wraps an OutOfMemoryError. 
Currently there is no way of enforcing Java exception type conventions during program development.
Hence, further investigation is needed on finding ways to help developers in dealing with this
 problem, either preventing odd wrappings or enabling the developer to
 better deal with them.
Furthermore, only some of the Android developers surveyed in this study were aware of the 
rebustness threats caused by cross-type wrappings as the one cited above. This calls for empirical studies on the actual usefulness of Java's hybrid exception model.

@@

%\emph{\textbf{On the null pointer problem.}}
\subsection{On the null pointer problem.}
The null reference was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake''~\cite{hoare2}.
In this sudy, the null references were, infact, responsible for several reported issues - providing further evidence to Hoare's statement. Moreover many of the survey respondents recognized that NullPointerExceptions was one of the main causes of application crashes since they can happen almost 
anyware in the application code, and to make thinks worse the life-cycle of Android apps on which objects are constantly recreated (i.e., Activity and Fragment classes) favor this kind of exception to happen added to the fact that many Android framework methods return null without making such
return type explicitly on the documentation.
Such observations emphasizes the need for solutions to avoid NullPointerExceptions, such as:
(i) lightweight intra-method null pointer analysis as supported by
Java 8 @Nullable annotations\footnote{Already supported by tools such as Eclipse, IntelliJ, Android Studio 0.5.5 (release Apr. 2014) to detect potential 
null pointer dereferences at compile time.};
(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha~\cite{nanda2009accurate};
or (iii) language designs which avoid null pointers, such 
as Monads~\cite{Walde95} (as used in functional languages for values that may not be available 
or computations that may fail) could improve the robustness of Java programs. 
Some of the Android developers mentioned how @Nullable annotations could helpful to deal with
NullPointerExceptions and consequently prevent app crashes caused by it.

%\emph{\textbf{Preventing uncaught exceptions}}
\subsection{Preventing uncaught exceptions}
In this study we could observe undocumented runtime exceptions thrown by third party code,
and even undocumented checked exception thrown by a JNI interface.
Such undocumented exceptions make it difficult, and most of the times infeasible
for the client code to protect against `unforeseen'' situations that may happen 
while calling a library code. In the survey-based study Android developers were asked about
ways to prevent application crashes which are mainly caused by uncaught exceptions.
Many of them emphasized the importance of handling exceptions to prevent crashes
but also mentioned how difficult is to handle every specific exception that can happen.
Several developers recognized to follow a more reactive behaviour against crashes:
they advocate the use of crash report tools, and once the crash happens for the first time,
its reported and the application can be changed to better cope with the exceptions that crashed the crashes, consequently preventing future similar crashes.

One may think that the solution for the uncaught exceptions may be to define a general handler, 
which is responsible for handling any exception that is not
adequately handled inside the applications. Although this 
solution may prevent  the system from abruptly crashing, as mentioned by some of the surveyed developers such a general handler will not have enough
contextual information to adequately handle the exception, 
beyond storing a message in a log file and restarting the application.
However, such a handler cannot replace a carefully designed exception 
handling policy~\cite{Robil00}, which requires
third-party documentation on the exceptions that
may the thrown by APIs used. 
Since  documenting runtime exceptions is a tedious and error prone
task, this calls for tool support to automate the extraction of runtime exceptions
from library code. Initial steps in this direction have been proposed
by van Doorn and Steegmans \cite{van2005combining}.

%\section{Threats to Validity}
%\label{sec:threats}
%\noindent

%\emph{\textbf{Threats to Validity of the Mining Study.}} 
\subsection{Mining Study - Threats to Validity }

\emph{Internal Validity.} We used a heuristics-based parser to mine
exceptions from issues.  Our parsing strategy was conservative by default; for
example, we only considered exception names using a fully qualified class name
as valid exception identifiers, while, in many cases, developers use the
exception name in issue description. Conservative parsing may minimize false
positives, which was our initial target, but also tends to increase false
negatives, which means that some cases may have not been identified as
exceptions or stack traces. Our limited manual inspection did not reveal such
cases. Moreover, in this study we manually mapped the concerns related
to exceptions. To ensure the quality of the analysis, we calculated the
 interrater agreement after three independent 
developers classified a randomly selected sample (of 25 exception 
types from the total of 100); the interater agreement was high (96\%).

@@

Another threat relates to the fact that parts of our analysis 
are based on the availability of stack traces on issues reported on Github and Googlecode projects. 
In using these datasets, we make an underlying assumption: the stack traces reported on issues are 
representative of valid crash information of the applications. 
One way to mitigate this threat would be to access to the full 
set of crash data per application. Although some services exist 
to collect crash data from mobile applications~\cite{BugSe14,BugSn14,Googl14,Acra14},
they do not provide open access to the crash reports of their client applications.
In our study, we mitigated this threat by manually inspecting
the source code associated to a subset of the reported exception stack traces.
This subset comprises the stack traces related to the main findings 
of the study (e.g., ``undocumented runtime and checked exceptions'',
and ``cross-type wrappings'').

\subsection{Limitations of the Survey-based Study.}

%\noindent
